Condition-Based Maintenance for Complex Systems\\ based on Current Component Status\\ and Bayesian Updating of Component Reliability

### 2016-10-22 cbm-rbd-paper.tex
change notation again

@@

@@

We consider a system with components of $k=1,\ldots,K$ different types;
for each type $k$, there are $n_k$ exchangeable components in the system.

@@

For each type $k$ component, we assume that the lifetime $T_{k,i}$ ($i=1,\ldots,n_k$, $k = 1, \ldots, K$)
is Weibull distributed with fixed shape parameter $\beta_k > 0$,
in short $T_{k,i} \mid \lambda_k \sim \wei(\beta_k,\lambda_k)$,
with density and cdf
\begin{align}
\label{eq:weibulldens}
f(t_{k,i} \mid \lambda_k) &= \frac{\beta_k}{\lambda_k} (t_{k,i})^{\beta_k-1} e^{-\frac{(t_{k,i})^{\beta_k}}{\lambda_k}}\,, \\
\label{eq:weibullcdf}
F(t_{k,i} \mid \lambda_k) &= 1 - e^{-\frac{(t_{k,i})^{\beta_k}}{\lambda_k}} = P(T_{k,i} \leq t_{k,i} \mid \lambda_k)\,,
\end{align}
where $\lambda_k > 0$ and $t > 0$.

@@

In place of the usual parametrization in terms of $a_k$ and $b_k$,
we use the parameters $\nkz > 1$ and $\ykz > 0$
(the upper index ${}\uz$ is used to indicate that these are prior parameters)
because they have a simple interpretation.
They are defined as
\begin{align}
\nkz &= a_k - 1 & &\text{ and}
&
\ykz &= b_k / \nkz,
\label{eq:abtony}
\end{align}
where $\ykz$ can be interpreted as the prior guess for the scale parameter $\lambda_k$,
as $\E[\lambda_k\mid\nkz,\ykz] = \ykz$.
Using \eqref{eq:lambdainterpret},
we can thus translate an expert's statement of expected component lifetime into a corresponding value for $\ykz$.
$\nkz$ can be seen as a prior strength or pseudocount,
this will become clear in the discussion of the update step below.

The parametrization in terms of $\nkz$ and $\ykz$ also clarifies the nature of the combination
of prior information and data through Bayes' rule.
In the conjugate setting,
applying Bayes' rule simply means that the prior parameters $\nkz$ and $\ykz$
are updated to posterior parameters, which we denote by $\nkn$ and $\ykn$, respectively.
Assume we observe $n_k = e_k + c_k$ component lifetimes,
where $e_k$ is the number of actual failure events,
and $c_k$ is the number of right-censored observations.
We denote the failure times by $t_{k,1}, \ldots, t_{k,e_k}$,
and the censoring times by $t^+_{k,1}, \ldots, t^+_{k,c_k}$,
and collect them in an observation vector $\vec{t}_k = (t_{k,1}, \ldots, t_{k,e_k}, t^+_{k,1}, \ldots, t^+_{k,c_k})$.
Then, the updated, posterior parameters are
\begin{align}
\nkn &= \nkz + e_k\,, 
&
\ykn &=  \frac{\nkz \ykz + \tautk}{\nkz + e_k}\,,
\label{eq:ig-update}
\end{align}
where $\tautk = \sum_{i=1}^{e_k} (t_{k,i})^\beta + \sum_{i=1}^{c_k} (t^+_{k,i})^\beta$.
The upper index ${}\un$ indicates that these are posterior parameters resulting from an update with $n_k$ observations,
where we leave out the index $k$ for $n_k$ in the upper index to increase legibility.

From the simple update rule \eqref{eq:ig-update}, we see that
$\ykn$ is a weighted average of the prior parameter $\ykz$ and the maximum likelihood (ML) estimator $\tautk/e_k$,
with weights proportional to $\nkz$ and $e_k$, respectively.
$\nkz$ can thus be interpreted as a prior strength or pseudocount,
indicating how much our prior guess should weigh against the $e_k$ observed failure events.
Furthermore, $\V[\lambda\mid\nkz,\ykz] = (\ykz)^2 / (1 - 1/\nkz)$;
for fixed $\ykz$, a higher $\nkz$ indicates thus 
that more probability mass is concentrated around $\ykz$.

Using \eqref{eq:abtony} and \eqref{eq:ig-update}, the posterior distribution over $\lambda_k$ is
\begin{align}
\lambda_k \mid \nkz, \ykz, \vec{t}_k \sim \ig(\nkz + e_k + 1, \nkz \ykz + \tautk)\,.
\label{eq:ig-update-alpha}
\end{align}
%where we have added the prior parameters $\nkz$ and $\ykz$ in notation
%to emphasize that the posterior is a synthesis of prior information and data.
As this posterior \eqref{eq:ig-update-alpha} can be defined in terms of
the updated, posterior parameters $\nkn$ and $\ykn$,
we may also write
\begin{align*}
f(\lambda_k \mid \nkz, \ykz, \vec{t}_k) &= f(\lambda_k \mid \nkn, \ykn)\,.
\end{align*}
The iterative nature of Bayesian inference means that we can take the updated,
posterior parameters as new prior parameters and update them again using new data.
Indeed, we will use this property to use the extra information about components
that accumulates during operation of the system:
At any time $\tnow > 0$,
all non-failed components correspond to a right censored observation $\tpnow$;
any failed components instead contribute a fully observed, non-censored lifetime,
and both types of observations can be used in the update step \eqref{eq:ig-update}.
Before we write this out formally in Section~\ref{sec:postpred},
we will first present how system reliability functions can be derived based on component models.
%To keep notation simple,
%we will below denote the parameters of the inverse Gamma distribution by $\nkz$ and $\ykz$,
%regardless of them expressing expert information alone,
%or stemming from the combination of expert information and test data.

@@

Here we describe how system reliability functions, given fixed component models,
can be efficiently calculated using the survival signature.
We will show in Section~***
how this method can be used to find the system reliability function $\Rsys(t)$ at current time $\tnow$ for a monitored system,
giving us the RUL/RLD distribution on which we base our adaptive maintenance policy.

We consider systems of arbitrary reliability block diagram layout,
consisting of components of $K$ different types,
where there are $n_k$ components of type $k$ in the system,
and the total number of components in the system is $\sum_{k=1}^K n_k = n$.
An example is given in Figure~***.
Note that this system cannot be described as a nesting of series and parallel subsystems.
%This system is observed until time $\tnow$,
%leading to censored observation of certain component lifetimes within the system.

In a system with $n$ components, the state of the components can be expressed by the state vector
$\vec{x} = (x_1,x_2,\ldots,x_n) \in \{0,1\}^n$,
with $x_i=1$ if the $i$th component functions and $x_i=0$ if not.
The structure function $\phi : \{0,1\}^n \rightarrow \{0,1\}$, defined for all possible $\vec{x}$,
takes  the value 1 if the system functions and 0 if the system does not function for state vector $\vec{x}$ \cite{BP75}.
Most real-life systems are coherent,
which means that $\phi(\vec{x})$ is non-decreasing in any of the components of $\vec{x}$,
so system functioning cannot be improved by worse performance of one or more of its components.
Furthermore, one can usually assume that $\phi(0, \ldots, 0) = 0$ and $\phi(1, \ldots, 1) = 1$.

The survival signature \cite{2012:survsign} is a summary of the structure function
for systems with $K$ groups of exchangeable components.
Denoted by $\Phi(l_1,\ldots,l_K)$, with $l_k=0,1,\ldots,n_k$ for $k=1,\ldots,K$,
it is defined as the probability for the event that the system functions
given that precisely $l_k$ of its $n_k$ components of type $k$ function, for each $k\in \{1,\ldots,K\}$.
Essentially, this creates a $K$-dimensional partition for the event $\Tsys > t$,
such that $\Rsys(t) = P(\Tsys > t)$ can be calculated using the law of total probability,
\begin{align}
P(\Tsys > t) &= \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} P(\Tsys > t \mid C^1_t = l_1,\ldots, C^K_t = l_K)
                                                                                  P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
             &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
             &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K P(C^k_t = l_k)\,,
\label{eq:sysrel-survsign}
\end{align}
where $P(C^k_t = l_k)$ is the (predictive) probability that exactly $l_k$ components of type $k$ function at time $t$,
and the last equality holds as we assume that components of different types are independent.
Note that for coherent systems, the survival signature $\Phi(l_1,\ldots,l_K)$ is non-decreasing in each $l_k$.

@@

\subsection{Component posterior predictive distributions}
\label{sec:postpred}

To calculate the current system reliability curve,
we thus need the probabilities $P(C^k_t = l_k)$,
which give the number of components that function at times $t > \tnow$.
In the Bayesian framework, these probabilities are given
by a so-called posterior predictive distribution that can be derived from the posterior over $\lambda$.

Denote by $\nkn$ and $\ykn$ the parameters obtained
by updating the prior parameters $\nkz$ and $\ykz$ using test data.
The corresponding distribution $f(\lambda_l \mid \nkn, \ykn)$
thus accounts for both expert knowledge and test data.
However, following our comments at the end of Section~\ref{sec:weibull},
the component models can be further updated
using information gained from the current run of the system until $\tnow$:
Any failure times of failed components,
and the right-censoring time $\tpnow$ for each non-failed component
can be used to update $\nkn$ and $\ykn$.
We denote the resulting parameter values by $\nknow$ and $\yknow$.
(It is of course possible to update $\nkz$ and $\ykz$ directly to $\nknow$ and $\yknow$ if no test data is available.)
%Furthermore, the fact that the components in the system that still function have reached the age $\tnow$
%must be used when calculating the posterior predictive distribution.

In analogue to the notation from Section~\ref{sec:weibull},
let $n_k = e_k^{\tnow} + c_k^{\tnow}$ be the number of type $k$ components in the monitored system,
where $e_k^{\tnow}$ is the number of components that have failed by $\tnow$,
and $c_k^{\tnow}$ is the number of components that have not failed by $\tnow$.
We can thus collect these observations, as of $\tnow$, in a vector
$\vec{t}_k^{\tnow} = (t_{k,1}, \ldots, t_{k,e_k}, \tpnow, \ldots, \tpnow)$,
containing $c_k^{\tnow}$ right-censored observations $\tpnow$.

>>>>

\newcommand{\nk}{n_k}
\newcommand{\nkp}{n_k'}
\newcommand{\yk}{y_k}
\newcommand{\ykp}{y_k'}

@@

\newcommand{\Rsysnow}{R^{(t_\text{now})}_\text{sys}}
\def\Phinow{\Phi^{(t_\text{now})}}

@@

We consider a system with components of $k=1,\ldots,K$ different types;
for each type $k$, there are $N_k$ exchangeable components in the system.

@@

For each type $k$ component, we assume that the lifetime $T_{k,i}$ ($i=1,\ldots,N_k$, $k = 1, \ldots, K$)
is Weibull distributed with fixed shape parameter $\beta_k > 0$,
in short $T_{k,i} \mid \lambda_k \sim \wei(\beta_k,\lambda_k)$,
with density and cdf
\begin{align}
\label{eq:weibulldens}
f(t_{k,i} \mid \lambda_k) &= \frac{\beta_k}{\lambda_k} (t_{k,i})^{\beta_k-1} e^{-\frac{(t_{k,i})^{\beta_k}}{\lambda_k}}\,, \\
\label{eq:weibullcdf}
F(t_{k,i} \mid \lambda_k) &= 1 - e^{-\frac{(t_{k,i})^{\beta_k}}{\lambda_k}} = P(T_{k,i} \leq t_{k,i} \mid \lambda_k)\,,
\end{align}
where $\lambda_k > 0$ and $t > 0$.

@@

In place of the usual parametrization in terms of $a_k$ and $b_k$,
we use the parameters $\nk > 1$ and $\yk > 0$
%(the upper index ${}\uz$ is used to indicate that these are prior parameters)
because they are more simple to interpret.
They are defined as
\begin{align}
\nk &= a_k - 1 & &\text{ and}
&
\yk &= b_k / \nk,
\label{eq:abtony}
\end{align}
where $\yk$ can be interpreted as the prior guess for the scale parameter $\lambda_k$,
as $\E[\lambda_k\mid\nk,\yk] = \yk$.
Using \eqref{eq:lambdainterpret},
we can thus translate an expert's statement of expected component lifetime into a corresponding value for $\yk$.
$\nk$ can be seen as a prior strength or pseudocount,
this will become clear in the discussion of the update step below.

The parametrization in terms of $\nk$ and $\yk$ also clarifies the nature of the combination
of prior information and data through Bayes' rule.
In the conjugate setting,
applying Bayes' rule simply means that the prior parameters $\nk$ and $\yk$
are updated to posterior parameters, which we denote by $\nkp$ and $\ykp$, respectively.
Assume we observe $N_k = e_k + c_k$ component lifetimes,
where $e_k$ is the number of actual failure events,
and $c_k$ is the number of right-censored observations.
We denote the failure times by $t_{k,1}, \ldots, t_{k,e_k}$,
and the censoring times by $t^+_{k,1}, \ldots, t^+_{k,c_k}$,
and collect them in an observation vector $\vec{t}_k = (t_{k,1}, \ldots, t_{k,e_k}, t^+_{k,1}, \ldots, t^+_{k,c_k})$.
Then, the updated, posterior parameters are
\begin{align}
\nkp &= \nk + e_k\,, 
&
\ykp &=  \frac{\nk \yk + \tautk}{\nk + e_k}\,,
\label{eq:ig-update}
\end{align}
where $\tautk = \sum_{i=1}^{e_k} (t_{k,i})^\beta + \sum_{i=1}^{c_k} (t^+_{k,i})^\beta$.
%The upper index ${}\un$ indicates that these are posterior parameters resulting from an update with $n_k$ observations,
%where we leave out the index $k$ for $n_k$ in the upper index to increase legibility.

From the simple update rule \eqref{eq:ig-update}, we see that
$\ykp$ is a weighted average of the prior parameter $\yk$ and the maximum likelihood (ML) estimator $\tautk/e_k$,
with weights proportional to $\nk$ and $e_k$, respectively.
$\nk$ can thus be interpreted as a prior strength or pseudocount,
indicating how much our prior guess should weigh against the $e_k$ observed failure events.
Furthermore, $\V[\lambda\mid\nk,\yk] = (\yk)^2 / (1 - 1/\nk)$;
for fixed $\yk$, a higher $\nk$ indicates thus 
that more probability mass is concentrated around $\yk$.

Using \eqref{eq:abtony} and \eqref{eq:ig-update}, the posterior distribution over $\lambda_k$ is
\begin{align}
\lambda_k \mid \nk, \yk, \vec{t}_k \sim \ig(\nk + e_k + 1, \nk \yk + \tautk)\,.
\label{eq:ig-update-alpha}
\end{align}
%where we have added the prior parameters $\nkz$ and $\ykz$ in notation
%to emphasize that the posterior is a synthesis of prior information and data.
As this posterior \eqref{eq:ig-update-alpha} can be defined in terms of
the updated, posterior parameters $\nkp$ and $\ykp$,
we may also write
\begin{align*}
f(\lambda_k \mid \nk, \yk, \vec{t}_k) &= f(\lambda_k \mid \nkp, \ykp)\,.
\end{align*}
The iterative nature of Bayesian inference means that we can take the updated,
posterior parameters as new prior parameters and update them again using new data.
Indeed, we will use this property to use the extra information about components
that accumulates during operation of the system:
At any time $\tnow > 0$,
all non-failed components correspond to a right censored observation $\tpnow$;
any failed components instead contribute a fully observed, non-censored lifetime,
and both types of observations can be used in the update step \eqref{eq:ig-update}.
Before we write this out formally in Section~\ref{sec:postpred},
we will first present how system reliability functions can be derived based on component models.
%To keep notation simple,
%we will below denote the parameters of the inverse Gamma distribution by $\nkz$ and $\ykz$,
%regardless of them expressing expert information alone,
%or stemming from the combination of expert information and test data.

@@

Here we describe how system reliability functions, given fixed component models,
can be efficiently calculated using the survival signature.
We will show in Section~***
how this method can be used to find the system reliability function $\Rsysnow(t)$ at current time $\tnow$ for a monitored system,
giving us the RUL/RLD distribution on which we base our adaptive maintenance policy.

We consider systems of arbitrary reliability block diagram layout,
consisting of components of $K$ different types,
where there are $N_k$ components of type $k$ in the system,
and the total number of components in the system is $\sum_{k=1}^K N_k = N$.
An example system reliability block diagram is given in Figure~***.
Note that this system cannot be described as a nesting of series and parallel subsystems.
%This system is observed until time $\tnow$,
%leading to censored observation of certain component lifetimes within the system.

In a system with $N$ components, the state of the components can be expressed by the state vector
$\vec{x} = (x_1,x_2,\ldots,x_N) \in \{0,1\}^N$,
with $x_i=1$ if the $i$th component functions and $x_i=0$ if not.
The structure function $\phi : \{0,1\}^N \rightarrow \{0,1\}$, defined for all possible $\vec{x}$,
takes the value 1 if the system functions and 0 if the system does not function for state vector $\vec{x}$ \citep{BP75}.
Most real-life systems are coherent,
which means that $\phi(\vec{x})$ is non-decreasing in any of the components of $\vec{x}$,
so system functioning cannot be improved by worse performance of one or more of its components.
Furthermore, one can usually assume that $\phi(0, \ldots, 0) = 0$ and $\phi(1, \ldots, 1) = 1$.

The survival signature \citep{2012:survsign} is a summary of the structure function
for systems with $K$ groups of exchangeable components.
Denoted by $\Phi(l_1,\ldots,l_K)$, with $l_k=0,1,\ldots,N_k$ for $k=1,\ldots,K$,
it is defined as the probability for the event that the system functions
given that precisely $l_k$ of its $N_k$ components of type $k$ function, for each $k\in \{1,\ldots,K\}$.
Essentially, this creates a $K$-dimensional partition for the event $\Tsys > t$,
such that $\Rsys(t) = P(\Tsys > t)$ can be calculated using the law of total probability,
\begin{align}
P(\Tsys > t) &= \sum_{l_1=0}^{N_1} \cdots \sum_{l_K=0}^{N_K} P(\Tsys > t \mid C^1_t = l_1,\ldots, C^K_t = l_K)
                                                                                  P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
             &= \sum_{l_1=0}^{N_1} \cdots \sum_{l_K=0}^{N_K} \Phi(l_1,\ldots,l_K) P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
             &= \sum_{l_1=0}^{N_1} \cdots \sum_{l_K=0}^{N_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K P(C^k_t = l_k)\,,
\label{eq:sysrel-survsign}
\end{align}
where $P(C^k_t = l_k)$ is the (predictive) probability that exactly $l_k$ components of type $k$ function at time $t$,
and the last equality holds as we assume that components of different types are independent.
Note that for coherent systems, the survival signature $\Phi(l_1,\ldots,l_K)$ is non-decreasing in each $l_k$.

@@

\subsection{Component posterior predictive distributions at $\tnow$}
\label{sec:postpred}

To calculate the current system reliability function $\Rsysnow(t)$,
we thus need, for each $k=1,\ldots, K$, the probabilities $P(C^k_t = l_k)$
for the number of type $k$ components that function at times $t > \tnow$.
In the Bayesian framework, these probabilities are given
by a so-called posterior predictive distribution that can be derived from the posterior over $\lambda$.

Denote by $\nkz$ and $\ykz$ the parameters reflecting the knowledge base at system start-up time $t=0$.
These could have been obtained by updating prior parameters $\nk$ and $\yk$ (reflecting expert knowledge)
to $\nkp$ and $\ykp$ using test data,
or could be taken directly equal to $\nk$ and $\yk$ if no test data is available.
In both cases, the distribution $f(\lambda_l \mid \nkz, \ykz)$
thus accounts for all knowledge on component type $k$ that is available at system start-up.

Following our comments at the end of Section~\ref{sec:weibull},
the component models can be further updated
using information gained from the current run of the system until $\tnow$:
Any failure times of failed components,
and the right-censoring time $\tpnow$ for each non-failed component
can be used to update $\nkz$ and $\ykz$.
We denote the resulting parameter values by $\nknow$ and $\yknow$.
%(It is of course possible to update $\nkz$ and $\ykz$ directly to $\nknow$ and $\yknow$ if no test data is available.)
%Furthermore, the fact that the components in the system that still function have reached the age $\tnow$
%must be used when calculating the posterior predictive distribution.

In analogue to the notation from Section~\ref{sec:weibull},
let $N_k = e_k^{\tnow} + c_k^{\tnow}$ be the number of type $k$ components in the monitored system,
where $e_k^{\tnow}$ is the number of components that have failed by $\tnow$,
and $c_k^{\tnow}$ is the number of components that have not failed by $\tnow$.
We can thus collect these observations, as of $\tnow$, in a vector
$\vec{t}_k^{\tnow} = (t_{k,1}, \ldots, t_{k,e_k}, \tpnow, \ldots, \tpnow)$,
containing $c_k^{\tnow}$ right-censored observations $\tpnow$.

$\Phinow$
